\documentclass[11pt,a4paper,openright]{amsbook}

%Fonts

\usepackage{libertine}
\usepackage{euler}
\usepackage[T1]{fontenc}

%Packages.

%\usepackage[parfill]{parskip}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows}
\usepackage[all]{xy}
\usepackage{mathrsfs}
\usepackage{a4wide}
\usepackage{enumerate}
\usepackage{hyperref}
%\usepackage{showkeys}
\usepackage{setspace}
%\usepackage{fancyhdr}

%Theorem styles.

\theoremstyle{plain}
\newtheorem{conj}{Conjecture}
\newtheorem{theorem}{Theorem}%[section]
\newtheorem{lemma}[theorem]{Lemma}%[section]
\newtheorem{proposition}[theorem]{Proposition}%[section]
\newtheorem{corollary}[theorem]{Corollary}%[section]
\newtheorem{claim}[theorem]{Claim}%[section]
\newtheorem{question}[theorem]{Question}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem*{thm}{Theorem}

\theoremstyle{definition}%[section]
\newtheorem{definition}[theorem]{Definition}%[section]
\newtheorem{example}[theorem]{Example}%[section]
\newtheorem{project}{Project}

\theoremstyle{remark}%[section]
\newtheorem{remark}[theorem]{Remark}%[section]

%Numbering and contents

%\numberwithin{theorem}{section}
\setcounter{tocdepth}{5}
\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex}

% Editting \Chapter

%Shorthand for hats.

\newcommand{\G}{\mathcal{G}}
\newcommand{\E}{\widehat{E}}
\newcommand{\X}{\widehat{X}}
\newcommand{\A}{\widehat{A}}



\newcommand{\ucong}{\rotatebox{90}{$\cong$}}
\newenvironment{chpt}{\newpage\normalfont\Large\bfseries \begin{center}}{\end{center}}
\title{Inverse Semigroups in Coarse Geometry.} 
\date{April 2013}
\author{Martin Finn-Sell}


\begin{document}
%\maketitle

\input{titlepage.tex}

%Line Spacing
\setlength{\oddsidemargin}{0.3in}
\setlength{\evensidemargin}{-0.4in}
\setlength{\textwidth}{15.4cm}

\bibliographystyle{alpha}
%\setstretch{1.3}

\frontmatter
\input{abstract.tex}


%
%\newpage
%\thispagestyle{empty}~
%\newpage
\addcontentsline{toc}{chapter}{Contents.}
\tableofcontents
\setstretch{1.3}
\begin{chpt}Author's Declaration.\end{chpt}
\addcontentsline{toc}{chapter}{Author's Declaration.}
I, Martin Finn-Sell, declare that the thesis entitled \textit{Inverse Semigroups in Coarse Geometry} and the work presented in the thesis are both my own, and have been generated by me as the result of my own original research. I confirm that:
\begin{itemize}
\item this work was done wholly or mainly while in candidature for a research degree at this University;
\item where any part of this thesis has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated;
\item where I have consulted the published work of others, this is always clearly attributed;
\item where I have quoted from the work of others, the source is always given. With the exception of such quotations, this thesis is entirely my own work;
\item I have acknowledged all main sources of help;
\item where the thesis is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself;
\item parts of this work have been published, as a preprint: \textit{Spaces of Graphs, Boundary Groupoids and the Coarse Baum-Connes Conjecture}, Martin Finn-Sell and Nick Wright, 2012\\
URL: http://arxiv.org/abs/1208.4237
\end{itemize}
\bigskip
Signed ..............................................................................
Date ..............................................................................
\newpage
\begin{chpt}Acknowledgements.\end{chpt}
\addcontentsline{toc}{chapter}{Acknowledgements.}

The two most important people I would like to thank are my supervisor, Nick Wright, for giving me a phenomenal amount of time and enthusiasm in the last three years, and my wife, Dr Sarah Finn-Sell for giving me endless amounts of support as well as nodding, smiling and reminding me that ``expander graphs do not embed into Hilbert space '' twice daily.  Without the support and encouragement they have both given me I would not be where I am today.

In addition to my supervisor there are many other people I owe thanks. Most chiefly amongst these, in an obvious but yet completely benign order, are Ana Khukhro (for turning coffee into theorems for 45 minutes every morning for 3 years), Joe Tait(o) (for beating me at go and reminding me as often as possible about Grothendieck), Martin Fluch (for being the life and soul and master of winter barbeques), Rufus Willett (who was exceptionally kind to me at conferences and for putting up with a barrage of bad ideas), Chris Cave (for every question that has caught me out), Tom Harris (for just loving K-theory in all its forms), Alex Bailey (for knowing heroic amounts of category theory), Raff Rainone (for being totally and utterly italian and working too hard), Connor Smythe (for telling me all about persistence and strata), David Matthews (for being very keen about graphs and badgers) and Yago Antolin (for being the most amazing male model that I know).

Without all of you I would not be as smart (not to mention insane) as I am today. Thank you all.

Beyond all the people above I would also like to thank the EPSRC for funding my doctoral studies and the University of Southampton for funding my fourth year of study.


\mainmatter
\input{chapter1.tex}
\input{chapter2.tex}
\input{chapter3.tex}
\input{chapter4.tex}
\input{chapter5.tex}

\bibliography{ref.bib}

\end{document}